Pure Spin Current Injection in Hydrogenated Graphene Structures

### 2017-10-21 mu/article/prb/article.tex
Making correcttions in figures of structures.

\begin{abstract}
We present a theoretical study of spin-velocity injection (SVI) of a pure spin
current (PSC) induced by  a linearly polarized light that impinges normally on
the surface of  two 50\% hydrogenated noncentrosymmetric two-dimensional (2D)
graphene structures. The first structure, hydrogenated at only one side,
labeled Up, also known as graphone, and the second, labelled Alt, is
25\% hydrogenated at both sides. The hydrogenation opens an energy gap in both
structures. We analyze two possibilities: in the first, the spin is fixed
along a chosen direction, and the resulting SVI is calculated;
in the second, we choose the SVI direction along
the surface plane, and calculate the resulting
spin orientation. This is done by changing the energy
$\hbar\omega$ and polarization angle $\alpha$  of the incoming  light. The
results are calculated within a full electronic band structure scheme using the
Density Functional Theory (DFT) in the Local Density Approximation (LDA). The
maxima of the spin-velocities are reached when $\hbar\omega=0.084$\,eV and
$\alpha=35^\circ$ for the Up structure, and $\hbar\omega=0.720$\,eV and
$\alpha=150^\circ$ for the Alt geometry. We find a speed of 668\,Km/s and
645\,Km/s for the Up and the Alt structures, respectively, when the spin points
perpendicularly to the surface. Also, the response is maximized by
fixing the spin-velocity direction 
along a high symmetry  axis, obtaining a speed of 688Km/s
with the spin pointing at $13^\circ$ from the surface normal, for the Up, 
and 906 Km/s and the spin pointing at $60^\circ$ from the surface normal, for
the Alt system. These speed values are of order of magnitude larger than those of bulk semiconductors,
such as CdSe and GaAs, thus making the hydrogenated graphene structures
excellent candidates for spintronics applications. 
\end{abstract}

@@

Spintronics is an emerging research field of electronics in which the
manipulation and transport of the electron spin in a solid state materials  is
central, adding a new degree of freedom to conventional charge
manipulation.\cite{wolfSC04,fabianAPS07} At present, there is an increasing
interest in attaining the same level of control over the transport of spin at
micro- or nano-scales, as it has been done for the flow of charge in typical
3D-bulk based electronic devices.\cite{awschalomNP2007} Several semiconductor
spintronics devices have been proposed \cite{majumdarAPL06,
dattaAPL90,gotteNat16,pershinPRB08}, and some of them require spin polarized
electrical current \cite{awschalomSSBM13} or pure spin current (PSC). One of
the difficulties in creating measurable  spin current and
development of PSC based semiconductor devices is the fact that the spin
relaxation time in semiconducting materials can be too short to enable the
spin transport, thus resulting in a non-observable spin
current.\cite{murakamiSc03} For PSC there is no net motion of charge; spin-up
electrons move in a given direction, while spin-down electrons travel in the
opposite one. This effect can be due to one-photon absorption of linearly
polarized light by a semiconductor, with filled valence bands and empty
conduction bands, illuminated by light with photon energy larger than the
energy gap. This phenomenon can be due to spin injection,\cite{malPRB03} Hall
Effects,\cite{sinovaPRB04} interference of two optical beams,\cite{bhatPRL00,
najmaiePRB03} or one photon absorption of linearly polarized
light\cite{bhatPRL05}. The last effect has been observed in gallium arsenide
(GaAs),\cite{zhaoPRL2006, stevensPRL03} aluminum-gallium arsenide
(AlGaAs),\cite{stevensPRL03} and Co$_2$FeSi.\cite{kimuraNGPAM12}

@@

To calculate the velocity of the spin injection
${\cal V}^{\mathrm{a}\mathrm{b}}(\omega)$ along the direction $\hat{\mathbf{a}}$, at
which the spin moves in a
polarized state along direction $\hat{\mathbf{b}}$,

@@

states, and we take $\epsilon\to 0$ at the end of the calculation. 
Since the 
spin-splitting of the valence ($v$) bands is very small, we neglect it throughout
this work,\cite{nastosPRB07} and then
$\rho_{vv'}({\mathbf k};t)= 

@@

k})$ being the energy of the electronic band $n$ and $m$ at point ${\mathbf k}$ in the
irreducible Brillouin zone (IBZ),  $|n{\mathbf k}\rangle$ is the Bloch state,
and $\omega_{nm}({\mathbf k})=\omega_{n}({\mathbf k})-\omega_{m}({\mathbf k})$.
Using $\mathcal{O} = \mathrm{Tr}(\hat{\rho}\hat{\mathcal{O}})$ for the
expectation value of an observable $\mathcal{O}$, where $\mathrm{Tr}$ denotes
the trace, we obtain

>>>>

\begin{abstract}
We present a theoretical study of spin-velocity injection (SVI) of a pure spin
current (PSC) induced by  a linearly polarized light that impinges normally on
the surface of  two 50\% hydrogenated noncentrosymmetric two-dimensional (2D)
graphene structures. The first structure, hydrogenated at only one side,
labeled Up, also known as graphone, and the second, labelled Alt, is 25\%
hydrogenated at both sides. The hydrogenation opens an energy gap in both
structures. We analyze two possibilities: in the first, the spin is fixed along
a chosen direction, and the resulting SVI is calculated; in the second, we
choose the SVI direction along the surface plane, and calculate the resulting
spin orientation. This is done by changing the energy $\hbar\omega$ and
polarization angle $\alpha$  of the incoming  light. The results are calculated
within a full electronic band structure scheme using the Density Functional
Theory (DFT) in the Local Density Approximation (LDA). The maxima of the spin-
velocities are reached when $\hbar\omega=0.084$\,eV and $\alpha=35^\circ$ for
the Up structure, and $\hbar\omega=0.720$\,eV and $\alpha=150^\circ$ for the
Alt geometry. We find a speed of 668\,Km/s and 645\,Km/s for the Up and the Alt
structures, respectively, when the spin points perpendicularly to the surface.
Also, the response is maximized by fixing the spin-velocity direction along a
high symmetry  axis, obtaining a speed of 688Km/s with the spin pointing at
$13^\circ$ from the surface normal, for the Up, and 906 Km/s and the spin
pointing at $60^\circ$ from the surface normal, for the Alt system. These speed
values are of order of magnitude larger than those of bulk semiconductors, such
as CdSe and GaAs, thus making the hydrogenated graphene structures excellent
candidates for spintronics applications.
\end{abstract}

@@

Spintronics is an emerging research field of electronics in which the
manipulation and transport of the electron spin in a solid state materials  is
central, adding a new degree of freedom to conventional charge
manipulation.\cite{wolfSC04,fabianAPS07} At present, there is an increasing
interest in attaining the same level of control over the transport of spin at
micro- or nano-scales, as it has been done for the flow of charge in typical
3D-bulk based electronic devices.\cite{awschalomNP2007} Several semiconductor
spintronics devices have been proposed \cite{majumdarAPL06,
dattaAPL90,gotteNat16,pershinPRB08}, and some of them require spin polarized
electrical current \cite{awschalomSSBM13} or pure spin current (PSC). One of
the difficulties in creating measurable  spin current and
development of PSC based semiconductor devices is the fact that the spin
relaxation time in semPiconducting materials can be too short to enable the
spin transport, thus resulting in a non-observable spin
current.\cite{murakamiSc03} For PSC there is no net motion of charge; spin-up
electrons move in a given direction, while spin-down electrons travel in the
opposite one. This effect can be due to one-photon absorption of linearly
polarized light by a semiconductor, with filled valence bands and empty
conduction bands, illuminated by light with photon energy larger than the
energy gap. This phenomenon can be due to spin injection,\cite{malPRB03} Hall
Effects,\cite{sinovaPRB04} interference of two optical beams,\cite{bhatPRL00,
najmaiePRB03} or one photon absorption of linearly polarized
light\cite{bhatPRL05}. The last effect has been observed in gallium arsenide
(GaAs),\cite{zhaoPRL2006, stevensPRL03} aluminum-gallium arsenide
(AlGaAs),\cite{stevensPRL03} and Co$_2$FeSi.\cite{kimuraNGPAM12}

@@

To calculate the velocity of the spin injection ${\cal
V}^{\mathrm{a}\mathrm{b}}(\omega)$ along the direction $\hat{\mathbf{a}}$, at
which the spin moves in a polarized state along direction $\hat{\mathbf{b}}$,

@@

states, and we take $\epsilon\to 0$ at the end of the calculation. Since the
spin-splitting of the valence ($v$) bands is very small, we neglect it
throughout this work,\cite{nastosPRB07} and then $\rho_{vv'}({\mathbf k};t)=

@@

k})$ being the energy of the electronic band $n$ and $m$ at point ${\mathbf k}$
in the irreducible Brillouin zone (IBZ),  $|n{\mathbf k}\rangle$ is the Bloch
state, and $\omega_{nm}({\mathbf k})=\omega_{n}({\mathbf
k})-\omega_{m}({\mathbf k})$. Using $\mathcal{O} =
\mathrm{Tr}(\hat{\rho}\hat{\mathcal{O}})$ for the expectation value of an
observable $\mathcal{O}$, where $\mathrm{Tr}$ denotes the trace, we obtain
